---
title: "World Happiness"
author: "Group 16"
date: "2021/6/30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(infer)
library(GGally)
library(plotly)
library(RColorBrewer)
```



```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
#Read "World Happiness" dataset into R.
setwd(dir = "D:\\STUDY\\GLASGOW\\Semester-2\\Data Analysis Skills\\Gruopwork\\Groupwork - 1 poster")
happy <- read_csv("happiness.csv")
happy <- as_tibble(happy)
```

This code adds a column to the data, so each country is either Africa or Rest of the World.



```{r addcol}
#Add a column called African, if countries are in Africa, the value is Africa, otherwise, the value is World.
happy <- happy %>%
  mutate(African = ifelse(happy$`Regional indicator` == "Sub-Saharan Africa" | happy$`Regional indicator` == "Middle East and North Africa",
                "Africa","World"))

```

# Introduction {#sec:Intro}

The World Happiness Dataset is the iconic statistical data of global happiness. It contains six critical factors: economic production, social support, life expectancy, freedom, absence of corruption, and generosity. To find which factors are most influential to the ladder score, we fitted a linear regression model with all relevant explanatory variables. Noting that Africa’s ladder score seems the smallest amongst all countries, we used bootstrapping to find the 95th percentile of the difference in ladder score between Africa and the rest of the world and fit the regression model for these two cases.

# Explanatory analysis {#sec:Explan}



```{r summ}
#Give summary to Happiness score, including n, mean, standard deviation, 1st quantile, 3rd quantile, min and max value, median value.
happy %>%
  group_by(`Regional indicator`) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25), TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`)) %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```



```{r}
#The graph shows the Happniess score in each region.
ggplot(data = happy,mapping = aes(y=`Regional indicator`,x=`Ladder score`,color=`Regional indicator`))+
  geom_point()+
  labs(x="Happiness score",y="Regions")+
  theme(legend.position="none")
```

Also summary data for Africa and Rest of the World.

```{r sumafrica}
#summary stats for AFRICA and REST OF WORLD
happy %>%
  group_by(African) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25),
            TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`))  %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")


```


This code uses bootstrapping to find a 95% confidence interval for the mean ladder score of all the data


```{r bootstrap}
bootstrap_distribution_mean <- happy %>% 
  specify(response = `Ladder score`) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean")

bootstrap_distribution_mean %>% 
  get_ci(level = 0.95, type = "percentile")

percentile_ci <- bootstrap_distribution_mean %>% 
  get_ci()

boots<-bootstrap_distribution_mean %>% 
  visualize(endpoints = percentile_ci, direction = "between")
png(filename = "boots.png",width=1920,height=1080,units="px",res=400)
boots

```


This next section of code produces a boxplot of ladder score, using region as a factor for all regions.

```{r box, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:scat} Ladder score by region.", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(`Regional indicator`), 
                                         y = `Ladder score`)) + geom_boxplot(fill = "slateblue") + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = unique(happy$`Regional indicator`))+
  theme(axis.text.x = element_text(angle=45,hjust = 0.5,vjust = 0.5))

```


Now same boxplot but just for Africa and rest of the World

```{r boxaf, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:bplot} Ladder score by Africa", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(African), 
                                   y = `Ladder score`,fill = "transparent"))+ 
  geom_boxplot(fill = brewer.pal(7, "Set1")[3]) + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = c("Africa","Rest of World"))

```


Using bootstrapping (week 7), we can find a 95th percentile of the difference in ladder score between Africa and rest of World.  
This shows that the score is between 1.65 and 1.11 lower in Africa than rest of World.

```{r boot}

bootstrap_distribution <- happy %>%
  specify(`Ladder score`~ African) %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in means", order = c("Africa", "World"))
percentile_ci <- bootstrap_distribution %>%
  get_ci()

percentile_ci

```

# Linear models {#sec:models}

## Question 1 

First we fit the full model includes six variables.

$$\widehat{\mbox{Ladder score}} = \widehat\alpha + \widehat\beta_{\mbox{1}} \cdot \mbox{GDP}
+\widehat\beta_{\mbox{2}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{3}} \cdot \mbox{life expectancy} + \widehat\beta_{\mbox{4}} \cdot \mbox{Freedom} + \widehat\beta_{\mbox{5}} \cdot \mbox{Generosity} + \widehat\beta_{\mbox{6}} \cdot \mbox{corruption}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,...,6);

```{r full_model}
full_model<-lm(`Ladder score`~`Logged GDP per capita`+`Social support`+`Healthy life expectancy`+`Freedom to make life choices`+`Generosity`+`Perceptions of corruption`,data = happy)
get_regression_table(full_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')

summary(full_model)
```

It shows that variables "Social support", "Freedom to make life choice" and "Logged GDP per capita" are most significant, then we check the correlationship between these six variables. 

```{r correlation}
ggpairs(happy[,c(3,7:12)])
```

Although "Healthy life expectancy" has the strongest correlation with "Ladder score", this variable also has the high correlation with other two variables (Logged GDP per capita and Social support). 

Finally, we decide to use three variables, Logged GDP per capita, Social support and Freedom to make life choices to fit our model.

$$\widehat{\mbox{Happiness score}} = \widehat\alpha 
+\widehat\beta_{\mbox{1}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{2}} \cdot \mbox{Freedom}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,2);

```{r final_model}
final_model<-lm(`Ladder score`~`Social support`+`Freedom to make life choices`+`Logged GDP per capita`,data = happy)
summary(final_model)
get_regression_table(final_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the final fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')
summary(final_model)
```

Our final model shows as follows,

$$\widehat{\mbox{Happiness score}} = -1.439
+5.649 \cdot \mbox{Social support} + 2.993 \cdot \mbox{Freedom}, $$

Before we can proceed to use the fitted model, we must check the assumption of the model. we assesses the model assumptions by plotting the residuals against the explanatory variable in Figure \ref{fig:resids_1} ,\ref{fig:resids_1} and a histogram (Figure \ref{fig:hist}) of the residuals to assess whether they are normally distributed.

```{r assumptions}
regression_points <- get_regression_points(final_model)
```

Check the first assumption of variable "Social support".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_1} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Social support`,y=residual)) +
  geom_point() +
  geom_jitter(width = 0.1) +
  labs(x = "Social support",y="Residual")+
  geom_hline(yintercept = 0,col="blue",size=1)
```

Check the first assumption of variable "Freedom to make life choices".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_2} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Freedom to make life choices`, y = residual)) +
  geom_point() +
  labs(x = "Freedom to make life choices", y = "Residual", title = "Residuals vs Freedom to make life choices")  +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```

Check the first assumption of variable "Logged GDP per capita".

```{r,echo=TRUE,fig.align='center',fig.width=13,fig.cap="\\label{fig:resids_2} Scatterplots of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points, aes(x = `Logged GDP per capita`, y = residual)) +
  geom_point() +
  labs(x = "Logged GDP per capita", y = "Residual", title = "Residuals vs Freedom to make life choices")  +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```

The scatterplots show an approximately spread of the residuals above and below the zero line, which means the data statisfied the first assumption, the residuals have mean zero.

```{r hist,echo=FALSE,fig.with=13,fig.align="center",fig.cap="\\label{fig:hist} Histgram of the residuals.",fig.pos='H',message=FALSE}
ggplot(regression_points,aes(x=residual))+
  geom_histogram(color="white")+
  labs(x="Residuals")
```

The histgram support the assumption of normally distributed errors in the model.

## Question 2

It is interesting to look at how corruption affects the happiness score for Africa compared to the rest of the World, and see if there are any differences.

Fitting an interaction model with separate intercepts and gradients.

$$\widehat{\mbox{Happiness score}} = \widehat\alpha 
+\widehat\beta_{\mbox{1}} \cdot \mbox{Region} + \widehat\beta_{\mbox{2}} \cdot \mbox{Corruption}+\widehat\beta_{\mbox{3}} \cdot \mbox{Region}\cdot \mbox{Corruption}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,2,3);

```{r}
#fit an interaction model
Afr.model <- lm(`Ladder score` ~ `Perceptions of corruption` * African, data = happy)
summary(Afr.model)
get_regression_table(Afr.model)
```

```{r}
#Draw the graph to show that corruption variable have different impact on Afircan countries and Non-African countries.
Q2<-ggplot(happy, aes(x = `Perceptions of corruption`, y = `Ladder score`, color = African)) +
  geom_jitter() +
  labs(x = "Perceptions of corruption", y = "Ladder Score", color = "African") +
  geom_smooth(method = "lm", se = FALSE)
png(filename = "Q2.png",width=1920,height=1080,units="px",res=400)
```
# Conclusions {#sec:Conc}
“Logged GDP per capita”, “Freedom to make life choice” and “Social support” have stronger linear relationship with “Happiness score”.

Most of African countries have lower happiness score than other countries.

Corruption has less impact on African countries, while for the rest of the world, this variables will significantly reduce happiness.
